### 2024-09-05 paper/header.tex
header, plots

% **************GENERATED FILE, DO NOT EDIT**************

\title{Digital Twins for Ocean Robots}

\author[1]{Gaël Forget}
\affil[1]{Earth, Atmospheric and Planetary Sciences, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, 02139, MA, USA}

\keywords{Julia, OptimizatOcean, Climate, Robot, Observation, Platform, Sensor, Drifter, Buoy, Profiler, Satellite, Modeling, Artificial Intelligence, Data Assimilation, Parameter Inference, Climatology, Geospatial, Statistics}

\hypersetup{
pdftitle = {Digital Twins for Ocean Robots},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Gaël Forget},
pdfkeywords = {Julia, OptimizatOcean, Climate, Robot, Observation, Platform, Sensor, Drifter, Buoy, Profiler, Satellite, Modeling, Artificial Intelligence, Data Assimilation, Parameter Inference, Climatology, Geospatial, Statistics},
}